\documentclass[12pt]{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage[bmargin=1in]{geometry}
\usepackage[inline]{asymptote}
\usepackage{comment}
\newenvironment{solution}
{\paragraph{Solution.}}
{\qed\eject}
\usepackage{hyperref}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\hyt}[2]{\hypertarget{#1 Claim #2}{\paragraph{Claim #2}}}
\newcommand{\hyl}[2]{Same as \hyperlink{#1 Claim #2}{Claim #2 in Solution #1}}
\newcommand{\hyr}[2]{\hyperlink{#1 Claim #2}{Claim #2}}
\DeclareMathOperator{\pow}{pow}
\newcommand{\nnt}{90^{\circ}}
\newcommand{\es}{\\[12pt]}
\usepackage{csquotes}
\usepackage{pythonhighlight}

\title{estimate}
\author{EGMOTC 2023 - Rohan}
\date{\today}
\setlength{\parindent}{0pt}

\begin{document}

\maketitle

\newcommand{\localtextbulletone}{\textcolor{black}{\raisebox{.45ex}{\rule{.6ex}{.6ex}}}}
\renewcommand{\labelitemi}{\localtextbulletone}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\thispagestyle{empty}

\section*{Categories}
\begin{itemize}
    \item I wrote some code!
    \item That's a large number!
    \item How many cats?
    \item I'm here for Math oly...
    \item Don't google that...
\end{itemize}

\eject

\section*{Problems}

\subsection*{I wrote some code!}

\begin{enumerate}
    \item [100] x=2023
    k=0
    while k<2023:
        x= (x/2) + (2023/(2*x))
        a=int(x/2)
        x=x+random.randint(1,a)
        k+=1
    print(int(x))
    
    \item [200] x=[]
    for i in range (10000):
        a=True
        l=list(str(i))
        for k in l:
            k=int(k)
            if k==0:
                a=False
            elif i%k!=0:
                a=False
        if a==True:
            x.append(i)
    print(len(x)) %339
    \item [400] 
    while k<23:
    for j in range(x):
        if j<=0:
            pass
        else:
            if x\%j==0:
                y+=random.randint(1,3*j)
            else:
                y+=random.randint(0,1)
    x=y
    y=0
    k+=1 
\end{enumerate}

\eject
\subsection*{That's a large number!}
\begin{itemize}
    \item [100] land mass of earth in sq. km = 510 million
    \item [200] power output of sun in joules per day ~ 3.3*10^31 J
    \item [400] how many pieces of art did picasso make? Answer: 147,800
\end{itemize}

\subsection*{Don't google that}
\begin{enumerate}
    \item [100] Number of volcano erruptions in 2023 ~72
    \item [200] views on EGMO23 slovenia video
    \item [400] Tallest tree ~ 379 feet
\end{enumerate}

\subsection*{Math Oly}
\begin{itemize}
    \item 
\end{itemize}

\end{document}